Tracing the Meta-Level: PyPy's Tracing JIT Compiler

### 2009-04-03 talk/icooolps2009/paper.tex
A minor simplification and add myself as an author

\numberofauthors{3}
\author{
\alignauthor Carl Friedrich Bolz\\
       \affaddr{Heinrich-Heine-Universität Düsseldorf}\\
       \affaddr{Softwaretechnik und Programmiersprachen}\\
       \affaddr{Institut für Informatik}\\ 
       \affaddr{Universitätsstra{\ss}e 1}\\
       \affaddr{D-40225 Düsseldorf}\\
       \affaddr{Deutschland}\\
       \email{cfbolz@gmx.de}
\alignauthor Antonio Cuni\\
       \affaddr{University of Genova}\\
       \affaddr{DISI}\\
       \affaddr{Italy}\\
       \email{cuni@disi.unige.it}
\alignauthor Armin Rigo\\
       \email{arigo@tunes.org}
}
\maketitle

@@

\fijal{Stopped reading at that point}
Let's look at which hints would need to be applied to the example interpreter
from Figure \ref{fig:tlr-basic}. The basic thing needed to apply hints is a
subclass of \texttt{JitDriver} that lists all the variables of the bytecode
loop. The variables are classified into two groups, red variables and green
variables. The green variables are those that the tracing JIT should consider to
be part of the program counter of the language interpreter. In the case of the
example, the \texttt{pc} variable is obviously part of the program counter.
However, the \texttt{bytecode} variable is also counted as green, since the
\texttt{pc} variable is meaningless without the knowledge of which bytecode
string is currently being interpreted. All other variables are red.

@@

The simple insight how to greatly improve the situation is that most of the
operations in the trace are actually concerned with manipulating the
bytecode and the program counter. Those are stored in variables that are part of
the position key (they are "green"), that means that the tracer checks that they
are some fixed value at the beginning of the loop. In the example the check
would be that the \texttt{bytecode} variable is the bytecode string
corresponding to the square function and that the \texttt{pc} variable is
\texttt{4}. Therefore it is possible to constant-fold computations on them away,
as long as the operations are side-effect free. Since strings are immutable in
Python, it is possible to constant-fold the \texttt{strgetitem} operation. The
\texttt{int\_add} operations can be folded anyway.

>>>>

\numberofauthors{3}
\author{
\alignauthor Carl Friedrich Bolz\\
       \affaddr{Heinrich-Heine-Universität Düsseldorf}\\
       \affaddr{Softwaretechnik und Programmiersprachen}\\
       \affaddr{Institut für Informatik}\\ 
       \affaddr{Universitätsstra{\ss}e 1}\\
       \affaddr{D-40225 Düsseldorf}\\
       \affaddr{Deutschland}\\
       \email{cfbolz@gmx.de}
\alignauthor Antonio Cuni\\
       \affaddr{University of Genova}\\
       \affaddr{DISI}\\
       \affaddr{Italy}\\
       \email{cuni@disi.unige.it}
\alignauthor Armin Rigo\\
       \email{arigo@tunes.org}
\alignauthor Maciej Fijalkowski\\
       \email{fijal@merlinux.eu}
}
\maketitle

@@

Let's look at which hints would need to be applied to the example interpreter
from Figure \ref{fig:tlr-basic}. The basic thing needed to apply hints is a
subclass of \texttt{JitDriver} that lists all the variables of the bytecode
loop. The variables are classified into two groups, red variables and green
variables. The green variables are those that the tracing JIT should consider to
be part of the program counter of the language interpreter. In the case of the
example, the \texttt{pc} variable is obviously part of the program counter.
However, the \texttt{bytecode} variable is also counted as green, since the
\texttt{pc} variable is meaningless without the knowledge of which bytecode
string is currently being interpreted. All other variables are red.

@@

The simple insight how to greatly improve the situation is that most of the
operations in the trace are actually concerned with manipulating the
bytecode and the program counter. Those are stored in variables that are part of
the position key (they are "green"), that means that the tracer checks that they
are some fixed value at the beginning of the loop. In the example the check
would be that the \texttt{bytecode} variable is the bytecode string
corresponding to the square function and that the \texttt{pc} variable is
\texttt{4}. Therefore it is possible to constant-fold computations on them away,
as long as the operations are side-effect free. Since strings are immutable in
Python, it is possible to constant-fold the \texttt{strgetitem} operation. The
\texttt{int\_add} are additions of constant \texttt{pc} and a true constant,
so they can be folded away.
